Using Metric Space Indexing for Complete and Efficient Record Linkage

### 2017-11-21 paper.tex

\usepackage{url}
\usepackage{hyperref}
\usepackage{cleveref}

@@

Systems that perform indexing prior to comparison and classification, as
illustrated in Fig.~\ref{fig-rl-process}, add a further practical
complexity to the process. Indexing, comparison and classification are
often conducted using algorithms and parameters selected based on
technical and domain expertise of the user of the record linkage system,
followed by a manual assessment (auditing) of the linkage
outcomes~\cite{Chr12}. If the quality of the resulting links is not good
enough for a certain application, the linkage process needs to be
repeated with different parameter settings and possibly also alternative
algorithms. This can result in a time-consuming and labor-intensive
iterative process. A major challenge is often the heuristic nature of
indexing, where the choice of a certain indexing technique and its
parameters (including which attributes to use in indexing) will
determine the final outcome of a linkage.

\textbf{TODO: GK rewrite this paragraph. Make the connection between completeness and high recall. Define completenes.}
Here we explore how existing indexing techniques can lead to
low recall, even for record pairs that are highly similar and likely
correspond to true matches. As a result, such techniques can lead
to \emph{incomplete} record linkage results. To address this problem,
\emph{metric space indexing} (MSI) can be used to guarantee that all
record pairs, up to a certain distance threshold, are considered in the
record linkage results. MSI also avoids the need for costly detailed
comparison of record pairs with low similarities, which may be inserted
into the same block with traditional indexing techniques. It allows
indexing, comparison and classification to be combined into a
single step, as shown in Fig.~\ref{fig-rl-process}, making the overall
record linkage process simpler, more efficient and more effective than
the traditional approach.

\textbf{Contribution:} The primary contribution of this paper is the novel application of MSI to achieve complete and efficient record linkage, without the need for complex parameter tuning.
We evaluate our approach on several data sets from diverse domains and demonstrate its advantages over
existing indexing techniques for record linkage.

% For final paper:
% The motivation for this work is the Digitising Scotland project (Dibben 2012), which is in the process of transcribing and linking all the vital events recorded in Scotland between 1856 and 1977. This data set will, when complete, include around 14 million birth records, 11 million death records and 4 million marriage records. As part of the work, certain data fields (locations, occupations and causes of death) are also being classified to the relevant standard coding schemes.

@@

@@

\smallskip
\textbf{Data Sets:}
%\label{sec-data}
We used four data sets from three domains in our experiments, as
summarized in Table~\ref{table-datasets}. The first is
\emph{CORA}~\footnote{Available from:
\texttt{http://secondstring.sourceforge.net}}, which contains 1,295
records that refer to 112 machine learning publications. 
It is used as a benchmark data set in the literature for assessing linkage algorithms. Ground truth is provided via a unique
\emph{paper\_id} identifier of the form "blum1993". In this experiment linkage is performed over the same set of records.

@@

\Cref{comparison-of-results} shows the results of the Cora experiments. The table shows the parameters for the experiment, the number of distance comparisons made, the precision, recall and F1 measure and the pairs quality and pair completeness achieved by each algorithm.

@@

\begin{figure}
\begin{verbatim}
This will be a table of Comparisons, F1, precision recall and pairs stuff.
12 ROWS ISH
\end{verbatim}
\caption{Linkage Quality on the Cora data set\label{comparison-of-results}}
\end{figure}

@@

\begin{figure}
\begin{verbatim}
This will be a table of Comparisons, F1, precision recall and pairs stuff for Skye/Kilmarnock
12 ROWS ISH
\end{verbatim}
\caption{Linkage Quality on the Cora data set\label{comparison-of-results}}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_Brute_Force}
\caption{Will be Linkage quality with M-Tree Skye\label{cora-block-quality-brute}}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_MTree}
\caption{Will be Linkage quality with M-Tree Kilmarnock\label{cora-block-quality-mtree}}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_MTree}
\caption{Will be Mtree F measure and a bunch LSH F measures\label{cora-block-quality-mtree}}
\end{figure}

@@

\section{Conclusions and Future Work}

\textbf{Contribution:} The primary contribution of this paper is the novel application of MSI to achieve complete and efficient record linkage, without the need for complex parameter tuning.
We evaluate our approach on several data sets from diverse domains and demonstrate its advantages over
existing indexing techniques for record linkage.

@@

\label{sec-concl}

@@

OLD STUFF

\section{The Plan}

\vspace{5mm}

Plan in a bulleted list:

\begin{itemize}
\item Design: Traditionally it has two stages. A blocking method needs to be developed separately from a similarity method. With similarity search based record linkage, there is only one method, and that is the similarity method.
\item Efficiency: I am not convinced of this, but Peter thinks there may be efficiency benefits. The reasoning behind this is that with blocking-based methods we would have to calculate some sort of a similarity value first. This value will only be used for blocking, and later a similarity value between candidate pairs will need to be calculated again.
\item Completeness: Similarity search based methods will be complete (with respect to the given similarity method) whereas blocking based methods (including LSH) are likely to be incomplete (there will be high-similarity pairs which are not places in the same block)
\item Data sets
\begin{itemize}
\item Kilmarnock Birth-Death
\item Skye Birth-Death
\item CORA
\item North Carolina Voter Registration DB
\end{itemize}
\item The following is a list of all methods we thought we would have to compare.
\begin{itemize}
\item Similarity search (M-Tree, others?)
\item LSH-blocking
\item Traditional blocking (e.g. Region, Surname) - \textbf{why this? - this was done in the paper: A Comparison of Blocking Methods for Record Linkage}
\item Sorted Neighbourhood blocking - \textbf{why this?}
\end{itemize}
\item Pairs-completeness and pairs-quality (w.r.t high similarity instead of true-link status). I will expand on this a bit more later.
\item Comparison (for blocking based methods)
		Precision/Recall/F1-Measure
\item Which similarity methods are metric?
\item Run time and memory usage comparisons
\item Theorem. Precision might get better or worse with similarity search based methods, but recall should never get worse.
\end{itemize}

Note: The rest of this paper should be regarded as a placeholder for now.

\smallskip
\textbf{MIfile:}
The MIFile \cite{amato2014mi} approach is an inexact technique for performing similarity search based on metric spaces. Using  MI\_file,  each record is represented by the ordering of distances from a collection of reference objects.
% Peter: are these reference object also records? 
It is based on the idea that two records close to each other will have similar neighbours and will thus be represented by similar of distances
% Peter: unclear what you mean with: similar of distances
from the chosen reference objects. 
In order to create the indexes, for each value to be added to the
% peter: value -> record ?
data-structure, first the k nearest reference objects (governed by a fixed parameter) are found, each is given a score (drawn from 1,2,3... based on the reference object's position from the value).
% peter: this is unclear .. I think I understand k-NN: for each
% record you find the k nearest reference objects and you enumerate
% them
Next for each score, a  mapping is created in an inverted index mapping
% peter: unclear..? a mapping in an inverted index mapping?
from the value to the score of each reference object and a representation of the data associated with the reference object and the value inserted.
% peter: unclear: what is "the data" associated with the reference
% object.
In order to perform a \textit{nearest-neighbour} lookup, first the k nearest reference objects (bounded by a second fixed parameter).
% peter: this sentence reads incomplete..?
Next the objects closest to these reference objects are extracted from the inverted map by choosing the neighbours with the highest score.
% Peter: sorry, this is unclear as well. "the objects closest" ->
% records closest? And why highest scores..? Is the closest record
% to a reference object given the highest score (not the lowest
% score - i.e. not 1-closest, 2-closest etc)
Much of the search space may be eliminated by  keeping track of the scores that  can contribute to the nearest neighbours.
% peter: so I guess you use the triangular inequality and the reference
% objects to prune the search space.

@@

%the following is a sentence we culled earlier but may be needed.
We will evaluate the performance of LSH based blocking with regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
Sect.~\ref{sec-exp}.

\begin{verbatim}
for each record a in A
    for each record b in B
        classify (a,b) as link if distance(a,b) < threshold
\end{verbatim}

% PAKDD 2018: maximum 12 pages
% Abstract max 200 words
% title, abstract: 1/2 page
% introduction: 1 1/2 page (2 pages so far)
% related work: 1 to 1 1/2 pages (3 pages so far)
% method: 4 pages (7 pages so far)
% experiments: 3 (10 pages so far)
% discussion /conclusion: 1 page
% citations 1

% Eamonn ICDM'10 tutorial slides
% - clear problem statement in abstract
% - To convince a reviewer, you must think like a reviewer

% Writing the paper:
% - Make a working title
% - Introduce the topic and define (informally at this stage)
%   terminology
% - Motivation: Emphasize why is the topic important
% - Relate to current knowledge: what’s been done
% - Indicate the gap: what need’s to be done?
% - Formally pose research questions
% - Explain any necessary background material.
% - Introduce formal definitions.
% - Introduce your novel algorithm/representation/data structure etc.
% - Describe experimental set-up, explain what the experiments will
%   show
% - Describe the datasets
% - Summarize results with figures/tables
% - Discuss results
% - Explain conflicting results, unexpected findings and discrepancies
%   with other research
% - State limitations of the study
% - State importance of findings
% - Announce directions for further research
% - Acknowledgements
% - References
%
% - Don’t make the reviewer of your paper think!
% - Reviewers make an initial impression on the first page and don’t
%   change 80% of the time
% - A good introduction with a good motivation is half your success
% - By the end of the introduction the reviewer mustknow.
%   - What is the problem?
%   - Why is it interesting and important?
%   - Why is it hard?why do naive approaches fail?
%   - Why hasn't it been solved before?(Or, what's wrong with previous
%     proposed solutions?)
%   - What are the key components of my approach and results?Also
%     include any specific limitations.
%   - A final paragraph or subsection: “Summary of Contributions”.
%     It should list the major contributions in bullet form,
%     mentioning in which sections they can be found. This material
%     doubles as an outline of the rest of the paper, saving space and
%     eliminating redundancy
% - Unjustified Choices (are bad)
% - Optimal: Does not mean `very good'
% - Proved: Does not mean `demonstrated'
% - Significant: There is a danger of confusing the informal statement
%   and the statistical claim
% - Use all the Space Available
% - Avoid Weak Language: aim, attempt, might, etc.
% - Use the Active Voice
% - ALWAYS put some variance estimate on performance measures (do
%   everything 10 times and give me the variance of whatever you are
%   reporting)

% Figures:
% - Don't cover the data with the labels!
% - Color helps -Direct labeling helps -Meaningful captions help
%
% Common problem with figures:
% 1.Too many patterns on bars
% 2.Use of both different symbols and different lines
% 3.Too many shades of gray on bars
% 4.Lines too thin (or thick)
% 5.Use of three-dimensional bars for only two variables
% 6.Lettering too small and font difficult to read
% 7.Symbols too small or difficult to distinguish
% 8.Redundant title printed on graph
% 9.Use of gray symbols or lines
% 10.Key outside the graph
% 11.Unnecessary numbers in the axis
% 12.Multiple colors map to the same shade of gray
% 13.Unnecessary shading in background
% 14.Using bitmap graphics (instead of vector graphics)
% 15.General carelessness

>>>>

\usepackage{url}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{hhline}

@@

Systems that perform indexing prior to comparison and classification, as
illustrated in Fig.~\ref{fig-rl-process}, add a further practical
complexity to the process. Indexing, comparison and classification are
often conducted using algorithms and parameters selected based on
technical and domain expertise of the user of the record linkage system,
followed by a manual assessment (auditing) of the linkage
outcomes~\cite{Chr12}. If the quality of the resulting links is not good
enough for a certain application, the linkage process needs to be
repeated with different parameter settings and possibly also alternative
algorithms. This can result in a time-consuming and labour-intensive
iterative process. A major challenge is often the heuristic nature of
indexing, where the choice of a certain indexing technique and its
parameters (including which attributes to use in indexing) will
determine the final outcome of a linkage.

This paper focuses on approaches that use a similarity threshold to
classify links. Such techniques are fundamentally limited by the extent
to which true matching records are similar, and true non-matches are
dissimilar---this is dataset dependent. Within this domain, we define
a technique to be \emph{complete} if it is guaranteed to identify all
pairs within the specified similarity threshold. Commonly used indexing
techniques are incomplete, since they reduce computational cost at the
expense of potentially overlooking some true matches. By definition,
incomplete techniques yield lower recall than complete techniques.
Conversely, and perhaps counter-intuitively, complete techniques can
yield lower precision with some datasets. This is discussed further in
Sect.~\ref{sec-approach}.

\emph{Metric space indexing} (MSI) is a complete technique with much
lower computational overheads than a brute-force approach. It allows
indexing, comparison and classification to be combined into a single
step, as shown in Fig.~\ref{fig-rl-process}, making the overall record
linkage process simpler, more efficient and more effective than
incomplete indexing approaches.

\textbf{Contribution:} the primary contribution of this paper is the
novel application of MSI to achieve complete and efficient record
linkage, without the need for complex parameter tuning. We evaluate our
approach on several data sets from diverse domains and demonstrate its
advantages over existing indexing techniques for record linkage.

% For final paper:
% The motivation for this work is the Digitising Scotland project
% (Dibben 2012), which is in the process of transcribing and linking all
% the vital events recorded in Scotland between 1856 and 1977. This data
% set will, when complete, include around 14 million birth records, 11
% million death records and 4 million marriage records. As part of the
% work, certain data fields (locations, occupations and causes of death)
% are also being classified to the relevant standard coding schemes.

@@

TODO: add stuff about how incomplete methods may yield higher precision than complete

@@

\smallskip
\textbf{Data Sets:}
%\label{sec-data}
We used four data sets from three domains in our experiments, as
summarized in Table~\ref{table-datasets}. The first is
\emph{Cora}~\cite{Cora2017}, which contains 1,295
records that refer to 112 machine learning publications. 
It is used as a benchmark data set in the literature for assessing linkage algorithms. Ground truth is provided via a unique
\emph{paper\_id} identifier of the form "blum1993". In this experiment linkage is performed over the same set of records.

@@

\Cref{comparison-of-results-cora} shows the results of the Cora experiments. The table shows the parameters for the experiment, the number of distance comparisons made, the precision, recall and F1 measure and the pairs quality and pair completeness achieved by each algorithm.  The linker labels may be interpreted as follows: the algorithm name is followed by its parameters, for LSH this is the size of the bands followed by the number of bands. For traditional blocking the parameters are the fields used for blocking.

@@

\begin{table}[ht]
\centering
\begin{tabular}{l|r|r|r|r|r|r}
Linker   & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\ \hline{=#=#=}
Brute Force        & 1,677,025                    & 0.02                  & 0.33                       & 0.84      & 0.57   & 0.68       \\ \hline
MTree              &  902,693                     & 0.89                  & 0.22                       & 0.84      & 0.57   & 0.68       \\ \hline
LSH-2-2            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-5-2            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-10-2           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-2-5            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-5-5            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-10-5           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-2-10           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-5-10           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-10-10          &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\  \hline
Blocking-year      &  115,893                     & 0.22                  & 0.19                       & 0.99      & 0.35   & 0.51       \\
Blocking-authors   &   11,039                      & 0.74                  & 0.1                        & 0.94      & 0.16   & 0.28       \\
Blocking-title     &   27,407                      & 0.84                  & 0.23                       & 0.95      & 0.42   & 0.58       \\
Blocking-venue     &   36,647                      & 0.7                   & 0.12                       & 0.85      & 0.29   & 0.44       \\
Blocking-location  & 1,009,957                    & 0.17                  & 0.24                       & 0.83      & 0.43   & 0.57       \\
Blocking-publisher &  833,079                     & 0.12                  & 0.23                       & 0.85      & 0.44   & 0.58       \\
Blocking-combined  & 1,214,269                    & 0.05                  & 0.32                       & 0.84      & 0.56   & 0.67       \\ \hline
\end{tabular}
\caption{Linkage Quality on the Cora data set distance=70\label{comparison-of-results-cora}
\textbf{TODO: The table needs some rows removed. Note that this table is with a threshold value of 70.}
}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{l|r|r|r|r|r|r}
Linker   & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\ \hhline{=#=#=}
MTree    & 102,318,525         & 0.84                  & 0.86                       & 0.65      & 0.46   & 0.54      \\ \hline
LSH-2-2  & 3,109,250           & 0                     & 0.05                       & 0.63      & 0.03   & 0.06       \\
LSH-5-2  & 10,412,496          & 0                     & 0.16                       & 0.64      & 0.11   & 0.19       \\
LSH-10-2 & 53,874,127          & 0                     & 0.46                       & 0.68      & 0.36   & 0.47       \\
LSH-5-5  & 36,566             & 0                     & 0.01                       & 0.76      & 0.01   & 0.01       \\
LSH-10-5 & 129,873            & 0                     & 0.01                       & 0.72      & 0.01   & 0.02       \\ \hline
\end{tabular}
\caption{Linkage Quality on demography datasets \textbf{This is for Skye. Threshold value is 2.}
\label{comparison-of-results-demography-skye}
}
\end{table}

\begin{table}[h]
\centering
\begin{tabular}{l|r|r|r|r|r|r}
Linker   & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\ \hline
MTree     & 514,871,153         & 0.89                  & 0.82                       & 0.76      & 0.45   & 0.57      \\ \hline
LSH-2-2   & 99,145,887          & 0                     & 0.23                       & 0.81      & 0.16   & 0.27       \\
LSH-5-2   & 130,721,338         & 0                     & 0.32                       & 0.79      & 0.23   & 0.36       \\
LSH-10-2  & 177,168,848         & 0                     & 0.48                       & 0.79      & 0.36   & 0.49       \\
LSH-5-5   & 239,368            & 0.01                  & 0.02                       & 0.84      & 0.01   & 0.02       \\
LSH-10-5  & 855,431            & 0.01                  & 0.03                       & 0.87      & 0.02   & 0.03 \\ \hline
\end{tabular}
\caption{Linkage Quality on demography datasets\textbf{This is for Kilmarnock. Threshold value is 2.
\label{comparison-of-results-demography-kili}
}
\end{table}

\begin{figure}
\includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_Brute_Force}
\caption{Linkage quality on Skye dataset using M-tree\label{skye-quality-mtree}}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_MTree}
\caption{Linkage quality on Kilmarnock dataset using M-tree\label{kilmarnock-quality-mtree}}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_MTree}
\caption{Comparison of quality attainable on Skye dataset using M-tree and various LSH configurations\label{mtree-lsh-quality}}
\end{figure}

@@

\section{Conclusions and Future Work\label{sec-concl}}

In this paper we have demonstrated the efficacy of MSI in achieving consistently complete and efficient record linkage, without the need for complex parameter tuning. In conclusion, this assertion deserves some careful unpacking. It is always possible to achieve high quality linkage using a brute-force approach. However the n-squared complexity of this approach prevents its practical application for datasets of even moderate size. The use of MSI techniques such as M-tree delivers the same results as the brute-force approach with fewer comparisons. We have shown that the M-tree approach delivers high precision, high recall results that are the same as those delivered by the brute-force approach. Furthermore this is achieved consistently without the need for complex parameter tuning.

We contrast this to traditional blocking and LSH based approaches. The major drawback of their use is that whilst they can produce extremely good results, they can also produce extremely poor results. It was our observations of low recall given by these approaches that originally led us to experiment with M-trees. Without extensive ground truth or large amounts of clerking effort by human domain experts it is impossible to determine the quality of the links produced.

However, this paper also highlights some unexpected results. Firstly, the good quality results yielded by both traditional blocking and LSH are partly due to the fact that (in the limit) they tend towards brute-force as the number of records in the block increase. A second unexpected result is that shown in \Cref{fig-cora-brute}, namely that LSH can yield higher precision than is achieved by a complete method such as M-Tree. This is due to complex reasons .... \emph{TODO EXPAND HERE}.

\emph{Some sentence or two about run-time?}

The observation that LSH can yield results quickly at the cost of some recall suggests a potentially profitable future direction for exploration, namely the use of hybrid algorithms. We hypothesise that it may be possible to use an incomplete technique like LSH to form indices into a data structure such as an M-tree. In this manner, M-tree nodes could be used as the starting point for range-search queries and, in so doing, prune the search space considerably to yield highly quality complete results quickly.

@@

@@

%text cut from earlier versions:

%The MIFile \cite{amato2014mi} approach is an inexact technique for performing similarity search based on metric spaces. Using  MI\_file,  each record is represented by the ordering of distances from a collection of reference objects.
% Peter: are these reference object also records? 
It is based on the idea that two records close to each other will have similar neighbours and will thus be represented by similar of distances
% Peter: unclear what you mean with: similar of distances
from the chosen reference objects. 
In order to create the indexes, for each value to be added to the
% peter: value -> record ?
data-structure, first the k nearest reference objects (governed by a fixed parameter) are found, each is given a score (drawn from 1,2,3... based on the reference object's position from the value).
% peter: this is unclear .. I think I understand k-NN: for each
% record you find the k nearest reference objects and you enumerate
% them
Next for each score, a  mapping is created in an inverted index mapping
% peter: unclear..? a mapping in an inverted index mapping?
from the value to the score of each reference object and a representation of the data associated with the reference object and the value inserted.
% peter: unclear: what is "the data" associated with the reference
% object.
In order to perform a \textit{nearest-neighbour} lookup, first the k nearest reference objects (bounded by a second fixed parameter).
% peter: this sentence reads incomplete..?
Next the objects closest to these reference objects are extracted from the inverted map by choosing the neighbours with the highest score.
% Peter: sorry, this is unclear as well. "the objects closest" ->
% records closest? And why highest scores..? Is the closest record
% to a reference object given the highest score (not the lowest
% score - i.e. not 1-closest, 2-closest etc)
Much of the search space may be eliminated by  keeping track of the scores that  can contribute to the nearest neighbours.
% peter: so I guess you use the triangular inequality and the reference
% objects to prune the search space.

@@

%We will evaluate the performance of LSH based blocking with regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
Sect.~\ref{sec-exp}.
